---
title: "Getting the data"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

- Using the `tidycensus` API to download the data.
- I had to install my [API key](http://api.census.gov/data/key_signup.html)
- That key gets installed in `.Renviron`. If it doesn't insall later, run `readRenviron("~/.Renviron")`.

I will handle the other tasks in another notebook.

```{r libraries, message=FALSE}
library(tidyverse)
library(tidycensus)
```


## Finding the variables

`tidycensus` has a `load_variables` [function](https://walkerke.github.io/tidycensus/articles/basic-usage.html#searching-for-variables) to help find and search for the fields you need.

This is commented out because I do not need it on subsequent runs once I found the variable I needed, which was `B01003_001`.

```{r}
# v17 <- load_variables(2017, "acs5", cache = TRUE)
# View(v17)
```

## Get the data

These are 5-year estimates from the 2013-2017 American Community Survey 5-Year Estimates table.

```{r}
tx_county_pop <- get_acs(geography = "county",
                         year = 2017,
                         variables = c(pop17 = "B01003_001"),
                         state = "TX")
head(tx_county_pop)
```

## Checking Harris County

Just to make sure I have [the right data](https://factfinder.census.gov/bkmk/cf/1.0/en/county/Harris County, Texas/POPULATION/ACS_EST).

```{r}
tx_county_pop %>% 
  filter(NAME == "Harris County, Texas")
```

## Save data to disc

This is so I don't have to download the data again to run analysis.

```{r}
saveRDS(tx_county_pop, "data/tx_county_pop.rds")
```

