---
title: "Making list, assign counties"
output: html_notebook
---

The purpose of this notebook is to rank the counties by population, then assign them to students to file PIRs.

- Rank the counties by population
- Give each student one county from the top of the list.
- Give each student an equal number of counties from the rest of the list.

```{r results="hide"}
library(tidyverse)
library(knitr)
```


## Load the data from disc

```{r}
tx_county_pop <- readRDS("data/tx_county_pop.rds")
head(tx_county_pop)
```

## Give counties a ranking

```{r}
tx_county_pop <- tx_county_pop %>% 
  arrange(desc(estimate)) %>% 
  mutate(rank = dense_rank(desc(estimate)))
tx_county_pop
```

## Make groups based on number of students

- Pull the `top_counties` based on the number of students. Each student would get one.
- Create a `small_counties` based on the remainder

```{r}
numb_students = 38 #number of students in the class

top_counties <- tx_county_pop %>% 
  arrange(rank) %>% 
  head(numb_students)

small_counties <- tx_county_pop %>%
  arrange(rank) %>% 
  tail(n = 254 - numb_students)
```

## Rank/select small counties

The idea here is to generate a random number for the small counties to give each student the same number of counties (or as close as possible).

```{r}

small_counties <- small_counties %>% 
  mutate(group_assigned=sample(rep(1:numb_students, length=nrow(small_counties)), replace=FALSE))
small_counties
```

### Check the samples 

```{r}
small_counties %>% 
  group_by(group_assigned) %>% 
  summarise(count = n())
```


### Order list and output

```{r}
small_counties %>% 
  group_by(group_assigned) %>% 
  arrange(group_assigned, NAME)

```

## Output data

```{r}

## save to disc
saveRDS(small_counties, "data/small_counties.rds")
saveRDS(top_counties, "data/top_counties.rds")

## write csv to disc
write_csv(small_counties, "data/small_counties.csv")
write_csv(top_counties, "data/top_counties.csv")
```

